---
title: "Reproducible Analysis of NCAA Volleyball BHE Change"
author: "John Forman"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: hide
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include = FALSE}
# Global chunk options and package load ---------------------------------------
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE
                      )

library(tidyverse)   # ggplot2, dplyr, tidyr, stringr, readr, purrr, etc.
library(dplyr)       # data manipulation
library(tidyr)       # data tidying
library(stringr)     # string manipulation
library(lubridate)   # date utilities
library(broom)       # convert model objects into tidy tibbles
library(fixest)      # DiD / fixed-effects models
library(ggplot2)     # data visualization
library(glue)        # string interpolation/helpers
library(scales)      # formatting axis scales (percent, comma, dollar, etc.)
library(knitr)       # simple tables for Word/PDF output
library(kableExtra)  # enhanced styling for tables in RMarkdown
```

```{r data-import}
# ---- 1. paths ---------------------------------------------------------------
# ⚠️ EDIT this path to your local project folder.
data_path <- "C:/Research/NCAA BHE Change/data/"

# ---- 2. read files ----------------------------------------------------------
matches <- read_csv(file.path(data_path, "analysis_ready_matches.csv"),
                    show_col_types = FALSE)
teams   <- read_csv(file.path(data_path, "analysis_ready_teams.csv"),
                    show_col_types = FALSE)
players <- read_csv(file.path(data_path, "analysis_ready_players.csv"),
                    show_col_types = FALSE)
```
## DESCRIPTIVES
# 1. Match-Level Overview (based on home team's division)
```{r d-stats1}
# Get each match’s home‐team division
home_division <- teams %>%
  filter(home_away == "Home") %>%
  select(match_id, division) %>%
  distinct()

# Join it into matches
matches <- matches %>%
  left_join(home_division, by = "match_id")

rm(home_division)


# Match-Level Overview (teams, matches, points, attendance)
match_summary <- matches %>%
  group_by(season, sex, division) %>%
  summarise(
    Total_Matches       = n(),
    Total_Teams         = n_distinct(home),
    Total_Points_Played = sum(total_points, na.rm = TRUE),
    Avg_Attendance      = mean(attendance, na.rm = TRUE),
    SD_Attendance       = sd(attendance, na.rm = TRUE),
    .groups = "drop"
  )

kable(match_summary, caption = "Match-Level Overview (based on home team's division)")

```

# 2. Ball-Handling Errors (BHE)
```{r d-stats2}
# Attach each match’s total_points into every team‐row
teams <- teams %>%
  left_join(
    matches %>% select(match_id, total_points),
    by = "match_id"
  )

# Summarize everything at the team‐level, grouping by division & home/away
summary_stats <- teams %>%
  group_by(season, sex, division, home_away) %>%
  summarise(
    Total_Matches   = n_distinct(match_id),
    Total_Points    = sum(total_points, na.rm = TRUE),
    Total_BHE       = sum(BHE, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Compute BHE per point
  mutate(
    BHE_per_Point = Total_BHE / Total_Points
  )

# Pivot so Home/Away become separate columns for each metric
final_table <- summary_stats %>%
  pivot_wider(
    names_from  = home_away,
    values_from = c(Total_Matches, Total_Points, Total_BHE, BHE_per_Point),
    names_glue  = "{.value}_{home_away}"
  )

kable(final_table, caption = "Match-Level Overview (based on home team's division)")

```

# 3. Position-Level Statistics (BHE by Setter vs Non-Setter)
```{r d-stats3}
# Bring total_points into players
players <- players %>%
  left_join(
    matches %>% select(match_id, total_points),
    by = "match_id"
  )

# For each match‐team and setter‐flag, collapse to one row per match:
player_match <- players %>%
  mutate(setter = (P == "S")) %>%
  group_by(match_id, sex, division, season, setter) %>%
  summarise(
    match_BHE     = sum(BHE, na.rm = TRUE),
    match_points  = unique(total_points),   # one value per match_id
    .groups       = "drop"
  )

# Now aggregate across all matches by sex/division/season/setter
tbl_setter <- player_match %>%
  group_by(sex, division, season, setter) %>%
  summarise(
    Total_BHE     = sum(match_BHE,    na.rm = TRUE),
    Total_Points  = sum(match_points, na.rm = TRUE),
    BHE_per_Point = Total_BHE / Total_Points,
    .groups       = "drop"
  )

kable(tbl_setter, caption = "Summarize BHE by Setter vs Non-Setter across division, sex, and season")
```


# 4. Rally Length (Attacks per Point)
```{r d-stats4}
# Aggregate total attacks per match
rally_match <- teams %>%
  group_by(match_id) %>%
  summarise(total_attacks = sum(TotalAttacks, na.rm = TRUE), .groups = 'drop') %>%
  left_join(
    matches %>% select(match_id, total_points, attendance, season),
    by = "match_id"
  ) %>%
  mutate(
    attacks_per_point = total_attacks / total_points,
    log_attendance    = log1p(attendance),
    post              = as.integer(season == "2024-25")
  )

# Basic table of mean attacks per point by season
rl_stats <- rally_match %>%
  group_by(season) %>%
  summarise(mean_attacks = mean(attacks_per_point, na.rm = TRUE), .groups = 'drop')
knitr::kable(rl_stats, digits = 3, caption = 'Mean attacks per point by season')

# Aggregate Total Attacks by sex, division, and season; derive attacks/rally and YoY % change
rally_length <- teams %>%
  group_by(season, division, sex) %>%
  summarise(
    # sum all attacks (home+away rows)
    Total_Attacks = sum(TotalAttacks, na.rm = TRUE),
    
    # sum each match’s points only once via home_away=="Home"
    Total_Points  = sum(
      if_else(home_away == "Home", total_points, 0),
      na.rm = TRUE
    ),
    
    # if you still want a count of matches:
    Total_Matches = n_distinct(match_id[home_away == "Home"]),
    .groups       = "drop"
  ) %>%
  mutate(
    Mean_Attacks_per_Point = Total_Attacks / Total_Points
  ) %>%
  arrange(division, sex, season) %>%
  group_by(division, sex) %>%
  mutate(
    Change_YoY = 100 * (Mean_Attacks_per_Point - lag(Mean_Attacks_per_Point))
    / lag(Mean_Attacks_per_Point)
  ) %>%
  ungroup()

kable(rally_length, caption = "Rally Length (Attacks per Point)")
```

# 5. Scoring Patterns (Distribution of Points Scored)
```{r d-stats5}
scoring_distribution <- teams %>%
# For each match, compute the opposing‐side errors
  group_by(match_id) %>%
  mutate(
    Opposing_Errors = 
      sum(BHE       [home_away != first(home_away)], na.rm = TRUE) +
      sum(SErr      [home_away != first(home_away)], na.rm = TRUE) +
      sum(BErr      [home_away != first(home_away)], na.rm = TRUE) +
      sum(Errors    [home_away != first(home_away)], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  
# Roll up to the season/division/sex level
  group_by(season, division, sex) %>%
  summarise(
    # raw totals
    Total_Kills   = sum(Kills,                na.rm = TRUE),
    Total_Aces    = sum(Aces,                 na.rm = TRUE),
    Total_Blocks  = sum(BlockSolos + BlockAssists/2, na.rm = TRUE),
    Total_Errors  = sum(Opposing_Errors,      na.rm = TRUE),
    # composite total
    Total_Points  = Total_Kills + Total_Aces + Total_Blocks + Total_Errors,
    # percentages
    Kills              = Total_Kills  / Total_Points * 100,
    Aces               = Total_Aces   / Total_Points * 100,
    Blocks             = Total_Blocks / Total_Points * 100,
    Opponent_Errors    = Total_Errors / Total_Points * 100,
    .groups = "drop"
  ) %>%
  arrange(division, sex, season)

kable(scoring_distribution, caption = "Distribution of Points Scored (as Percent of Total Points)")
```

# Table 1
```{r table-1}
# 1. Summarise by division/sex/season
tbl1 <- teams %>%
  group_by(division, sex, season) %>%
  summarise(
    # number of distinct teams
    n           = n_distinct(Name),
    # number of matches (count only the home side once)
    M           = n_distinct(match_id[home_away == "Home"]),
    # attendance mean/sd from the home rows
    mean_att    = mean(attendance[home_away == "Home"], na.rm = TRUE),
    sd_att      = sd(attendance[home_away == "Home"],   na.rm = TRUE),
    # total points from the home rows (total_points already in teams)
    total_pts   = sum(total_points[home_away == "Home"], na.rm = TRUE),
    # BHE counts
    bhe_home_ct = sum(BHE[home_away == "Home"],         na.rm = TRUE),
    bhe_away_ct = sum(BHE[home_away == "Away"],         na.rm = TRUE),
    .groups     = "drop"
  ) %>%
  mutate(
    bhe_home_rate = bhe_home_ct / total_pts,
    bhe_away_rate = bhe_away_ct / total_pts,
    # create the grouping label
    group = if_else(sex == "W", "Women", "Men†"),
    # roman numeral divisions
    Div = case_when(
      sex=="W" & division=="D1" ~ "I",
      sex=="W" & division=="D2" ~ "II",
      sex=="W" & division=="D3" ~ "III",
      sex=="M" & division=="D1" ~ "I",
      sex=="M" & division=="D3" ~ "III"
    ),
    # ordering key
    group_order = case_when(
      sex=="W" & division=="D1" ~ 1,
      sex=="W" & division=="D2" ~ 2,
      sex=="W" & division=="D3" ~ 3,
      sex=="M" & division=="D1" ~ 4,
      sex=="M" & division=="D3" ~ 5
    ),
    # formatted display columns
    `Mean Attend (SD)`  = paste0(comma(round(mean_att)),  " (", comma(round(sd_att)),  ")"),
    `Total Pts Played`  = comma(total_pts),
    `BHE Home (BHE/Pt)` = paste0(comma(bhe_home_ct), " (", sprintf("%.4f", bhe_home_rate), ")"),
    `BHE Away (BHE/Pt)` = paste0(comma(bhe_away_ct), " (", sprintf("%.4f", bhe_away_rate), ")")
  ) %>%
  arrange(group_order, season) %>%
  select(
    group, Div, n, season, M,
    `Mean Attend (SD)`,
    `Total Pts Played`,
    `BHE Home (BHE/Pt)`,
    `BHE Away (BHE/Pt)`
  ) %>%
  rename(` ` = group)

kable(
  tbl1,
  booktabs = TRUE,
  caption  = "Table 1. Team and Match Characteristics by Season, Division, and Sex."
) %>%
  kable_styling(full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "top")
```

# Table 2
```{r table-2-panelA}
# 1. Build per-match, per-setter BHE & points
match_setter <- players %>%
  mutate(setter = (P == "S")) %>%
  group_by(match_id, sex, division, season, setter) %>%
  summarise(
    match_BHE    = sum(BHE, na.rm = TRUE),
    match_points = unique(total_points),
    .groups      = "drop"
  )

# 2. Roll up to (sex,division,season,setter) & compute rate
tbl2a_raw <- match_setter %>%
  group_by(sex, division, season, setter) %>%
  summarise(
    Total_BHE    = sum(match_BHE,    na.rm = TRUE),
    Total_Points = sum(match_points, na.rm = TRUE),
    .groups      = "drop"
  ) %>%
  mutate(
    BHE_per_Point = Total_BHE / Total_Points
  )

# 3. Bring in rally‐length metrics (Total_Attacks & Mean-per-Point + YoY)
tbl2a <- tbl2a_raw %>%
  left_join(
    rally_length %>% select(sex, division, season, Total_Attacks, Mean_Attacks_per_Point, Change_YoY),
    by = c("sex","division","season")
  ) %>%
  
  # 4. Pivot setter → wide columns
  pivot_wider(
    names_from  = setter,
    values_from = c(Total_BHE, Total_Points, BHE_per_Point),
    names_glue  = "{.value}_{ifelse(setter,'Setter','NonSetter')}"
  ) %>%
  
  # 5. Compute the “total” BHE/Pt as sum of the two rates
  mutate(
    Total_BHE_Pt = BHE_per_Point_Setter + BHE_per_Point_NonSetter
  ) %>%
  
  # 6. Prepare formatting strings & YoY annotations
  group_by(sex, division) %>%
  arrange(sex, division, season) %>%
  mutate(
    # setter & non-setter rates with YoY on 2nd year
    Setter_fmt     = sprintf("%.4f", BHE_per_Point_Setter),
    Setter_YoY     = if_else(season=="2024-25",
                             paste0(" (", round(100*(BHE_per_Point_Setter - lag(BHE_per_Point_Setter)) 
                                               /lag(BHE_per_Point_Setter),1), "%)"),
                             ""),
    NonSetter_fmt  = sprintf("%.4f", BHE_per_Point_NonSetter),
    NonSetter_YoY  = if_else(season=="2024-25",
                             paste0(" (", round(100*(BHE_per_Point_NonSetter - lag(BHE_per_Point_NonSetter)) 
                                               /lag(BHE_per_Point_NonSetter),1), "%)"),
                             ""),
    # total BHE/Pt
    Total_fmt      = sprintf("%.4f", Total_BHE_Pt),
    # Attacks/Rally with YoY annotation
    AR_fmt         = paste0(
                       sprintf("%.3f", Mean_Attacks_per_Point),
                       if_else(is.na(Change_YoY),
                               "",
                               paste0(" (", round(Change_YoY,1), "%)")
                       )
                     )
  ) %>%
  ungroup() %>%
  
  # 7. Add grouping labels & Roman Div, order blocks
  mutate(
    group    = if_else(sex=="W","Women","Men†"),
    Div      = case_when(
                 sex=="W" & division=="D1" ~ "I",
                 sex=="W" & division=="D2" ~ "II",
                 sex=="W" & division=="D3" ~ "III",
                 sex=="M" & division=="D1" ~ "I",
                 sex=="M" & division=="D3" ~ "III"
               ),
    grp_ord  = case_when(
                 sex=="W" & division=="D1" ~ 1,
                 sex=="W" & division=="D2" ~ 2,
                 sex=="W" & division=="D3" ~ 3,
                 sex=="M" & division=="D1" ~ 4,
                 sex=="M" & division=="D3" ~ 5
               )
  ) %>%
  arrange(grp_ord, season) %>%
  
  # 8. Select & rename into final display
  transmute(
    ` `                           = group,
    Div,
    Season                        = season,
    `Setter BHE/Pt (YoY % change)`    = paste0(Setter_fmt, Setter_YoY),
    `Non-Setter BHE/Pt (YoY % change)`= paste0(NonSetter_fmt, NonSetter_YoY),
    `Total BHE/Pt`                = Total_fmt,
    `Total Attacks`               = comma(Total_Attacks),
    `Attacks/Rally (YoY % change)`   = AR_fmt
  )

# 9. Render with kableExtra, collapsing the “Women”/“Men†” label
kable(
  tbl2a,
  booktabs = TRUE,
  caption  = "Table 2, Panel A: Impact on setters and rally length."
) %>%
  kable_styling(full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r table-2-panelB}
# 1. Compute opposing‐side errors for each team‐row
team_errors <- teams %>%
  group_by(match_id) %>%
  mutate(
    Opposing_Errors = sum(
      BHE   [home_away != first(home_away)] +
      SErr  [home_away != first(home_away)] +
      BErr  [home_away != first(home_away)] +
      Errors[home_away != first(home_away)],
      na.rm = TRUE
    )
  ) %>% 
  ungroup()

# 2. Summarise raw totals by sex/division/season
dist_raw <- team_errors %>%
  group_by(sex, division, season) %>%
  summarise(
    Total_Kills   = sum(Kills,                         na.rm = TRUE),
    Total_Aces    = sum(Aces,                          na.rm = TRUE),
    Total_Blocks  = sum(BlockSolos + BlockAssists/2,   na.rm = TRUE),
    Total_Errors  = sum(Opposing_Errors,               na.rm = TRUE),
    .groups       = "drop"
  ) %>%
  mutate(
    Total_Pts   = Total_Kills + Total_Aces + Total_Blocks + Total_Errors,
    kills_prop  = Total_Kills  / Total_Pts,
    aces_prop   = Total_Aces   / Total_Pts,
    blocks_prop = Total_Blocks / Total_Pts,
    errs_prop   = Total_Errors / Total_Pts
  )

# 3. Format percentages with YoY for 2024-25
dist_fmt <- dist_raw %>%
  group_by(sex, division) %>%
  arrange(season) %>%
  mutate(
    kills_fmt  = paste0(
                    percent(kills_prop,  accuracy = 0.01),
                    if_else(season=="2024-25",
                            paste0(" (", round(100*(kills_prop - lag(kills_prop))/lag(kills_prop),1), "%)"),
                            "")
                  ),
    aces_fmt   = paste0(
                    percent(aces_prop,   accuracy = 0.01),
                    if_else(season=="2024-25",
                            paste0(" (", round(100*(aces_prop - lag(aces_prop))/lag(aces_prop),1), "%)"),
                            "")
                  ),
    blocks_fmt = paste0(
                    percent(blocks_prop, accuracy = 0.01),
                    if_else(season=="2024-25",
                            paste0(" (", round(100*(blocks_prop - lag(blocks_prop))/lag(blocks_prop),1), "%)"),
                            "")
                  ),
    errs_fmt   = paste0(
                    percent(errs_prop,   accuracy = 0.01),
                    if_else(season=="2024-25",
                            paste0(" (", round(100*(errs_prop - lag(errs_prop))/lag(errs_prop),1), "%)"),
                            "")
                  )
  ) %>%
  ungroup() %>%
  
  # Add grouping labels & Roman Div + ordering key
  mutate(
    group   = if_else(sex=="W", "Women", "Men†"),
    Div     = case_when(
                sex=="W" & division=="D1" ~ "I",
                sex=="W" & division=="D2" ~ "II",
                sex=="W" & division=="D3" ~ "III",
                sex=="M" & division=="D1" ~ "I",
                sex=="M" & division=="D3" ~ "III"
              ),
    grp_ord = case_when(
                sex=="W" & division=="D1" ~ 1,
                sex=="W" & division=="D2" ~ 2,
                sex=="W" & division=="D3" ~ 3,
                sex=="M" & division=="D1" ~ 4,
                sex=="M" & division=="D3" ~ 5
              )
  ) %>%
  arrange(grp_ord, season) %>%
  
  # 4. Select & rename into final display columns
  transmute(
    ` `                        = group,
    Div,
    Season                     = season,
    `Kills (YoY % gain)`       = kills_fmt,
    `Aces (YoY % gain)`        = aces_fmt,
    `Blocks (YoY % gain)`      = blocks_fmt,
    `Opponent Errors (YoY % gain)` = errs_fmt
  )

# 5. Render with collapsed “Women”/“Men†” label
kable(
  dist_fmt,
  booktabs = TRUE,
  caption  = "Table 2, Panel B: Impact on the distribution of points won"
) %>%
  kable_styling(full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "top")
```

# Interrupted Time‑Series models {-}
## Model results
```{r its-model}
# Run the models and show the summary results
its_women <- feols(BHE_per_point ~ time_index + post + time_index:post + log_attendance |
                     Name, cluster = ~ Name, data = filter(teams, sex == "W"))
its_men   <- feols(BHE_per_point ~ time_index + post + time_index:post + log_attendance |
                     Name, cluster = ~ Name, data = filter(teams, sex == "M"))

list(Women = summary(its_women), Men = summary(its_men))
```

## Table 3
```{r table-3}
# 1. Get sample sizes for the headers
sizes <- teams %>%
  group_by(sex) %>%
  summarise(
    n_matches = n_distinct(match_id),
    n_teams   = n_distinct(Name),
    .groups   = "drop"
  )
women_sizes <- sizes %>% filter(sex=="W")
men_sizes   <- sizes %>% filter(sex=="M")
women_lbl   <- glue("Women (n={comma(women_sizes$n_matches)}; teams={comma(women_sizes$n_teams)})")
men_lbl     <- glue("Men   (n={comma(men_sizes$n_matches)}; teams={comma(men_sizes$n_teams)})")

# 2. Tidy up each model
tw <- tidy(its_women, conf.int=FALSE) %>%
  select(term, estimate, std.error) %>%
  rename(Est_W = estimate, SE_W = std.error)

tm <- tidy(its_men, conf.int=FALSE) %>%
  select(term, estimate, std.error) %>%
  rename(Est_M = estimate, SE_M = std.error)

# 3. Combine and format into a display table
tbl3 <- tw %>%
  inner_join(tm, by="term") %>%
  # keep only the four rows you care about, in order
  filter(term %in% c("time_index","post","time_index:post","log_attendance")) %>%
  mutate(
    ` `      = c("Time index", "Post-rule", "Time × Post", "log(Attendance)"),
    Est_W    = sprintf("%.4f", Est_W),
    SE_W     = sprintf("(%.4f)", SE_W),
    Est_M    = sprintf("%.4f", Est_M),
    SE_M     = sprintf("(%.4f)", SE_M)
  ) %>%
  select(` `, Est_W, SE_W, Est_M, SE_M)

# 4. Render with kableExtra, adding the top‐spanning header
kable(
  tbl3,
  booktabs = TRUE,
  col.names = c("", "β", "SE", "β", "SE"),
  caption   = "Table 3. Interrupted time‐series estimates for BHE per point."
) %>%
  add_header_above(c(
    " "        = 1,
    setNames(2, women_lbl),
    setNames(2, men_lbl)
  )) %>%
  kable_styling(full_width = FALSE)
```

## Pre/Post BHE Trends by Sex & Season plot (Appendix Figure A1)
```{r its-plot, fig.cap='Weekly mean BHE per point (solid) with segmented OLS fit (dotted)', echo=FALSE}
# 1. Aggregate BHE per point by calendar week --------------------------------
bhe_week <- teams %>%
  mutate(week = floor_date(match_date, unit = "week", week_start = 1)) %>%
  group_by(sex, week) %>%
  summarise(BHE = mean(BHE_per_point, na.rm = TRUE), .groups = "drop")

# 2. Create prediction data from ITS models -----------------------------
first_date <- min(matches$match_date, na.rm = TRUE)

# pick one existing team name per sex (for dummy FE column)
dummy_names <- teams %>% distinct(sex, Name) %>% group_by(sex) %>% slice(1) %>% ungroup()

pred_grid <- bhe_week %>%
  mutate(time_index = as.numeric(difftime(week, first_date, units = "days")),
         post       = as.integer(week >= as.Date("2024-08-01")),
         log_attendance = 0) %>%              # hold attendance at 0
  left_join(dummy_names, by = "sex")         # supplies 'Name' column

pred_w <- pred_grid %>%
  filter(sex == "W") %>%
  mutate(pred = predict(its_women, newdata = ., fixef = FALSE))

pred_m <- pred_grid %>%
  filter(sex == "M") %>%
  mutate(pred = predict(its_men, newdata = ., fixef = FALSE))

pred_all <- bind_rows(pred_w, pred_m)

# 3. Plot --------------------------------------------------------------------
ggplot() +
  geom_line(data = bhe_week, aes(week, BHE, colour = sex), size = 0.6) +
  geom_line(data = pred_all, aes(week, pred, colour = sex), linetype = "dotted", size = 0.8) +
  geom_vline(xintercept = as.Date("2024-08-01"), linetype = "dashed") +
  scale_colour_manual(values = c("M" = "#0072B2", "W" = "#D55E00"), labels = c("Men", "Women")) +
  scale_y_continuous(labels = label_number(accuracy = 0.0001)) +
  labs(x = "Week (Monday start)", y = "BHE per point", colour = "Sex", title   = "Pre/Post BHE Trends by Sex & Season") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top")
```

# Position-based Difference-in-Differences (Setter vs Other) {-}
## Model results

```{r setter-did}
# 1. Create df with aggregated rows for setters and non-setters for each team each match
setter_match <- players %>%
   # flag Setters vs non-Setters
   mutate(setter = (P == "S")) %>%
   
  # aggregate at match × team (Name) × setter-status
  group_by(match_id, season, Name, setter) %>%
  summarise(
    total_BHE = sum(BHE, na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  
   # bring in log_attendance, home_away & total_points from teams
  left_join(
    teams %>%
      select(match_id, Name, home_away, total_points, log_attendance),
    by = c("match_id", "Name")
  ) %>%
  
   # compute per-point rate & post indicator
  mutate(
    BHE_per_point = total_BHE / total_points,
    post          = as.integer(season == "2024-25")
  )


# 2. Run DiD model: fixed effects for match + season -------------------------
setter_did <- feols(BHE_per_point ~ setter * post + home_away + log_attendance |
                     match_id + season,
                   cluster = ~ match_id,
                   data = setter_match)

summary(setter_did)
```

## Table 4
```{r table-4}
# 1. Tidy the DiD model, extracting clustered SEs, 95% CIs, and p‐values
tbl4 <- tidy(setter_did, conf.int = TRUE) %>%
  # keep only the three terms you care about
  filter(term %in% c("setterTRUE", "home_awayHome", "setterTRUE:post")) %>%
  
# 2. Relabel and format
  transmute(
    Predictor        = case_when(
                         term == "setterTRUE"     ~ "Setter (1 / 0)",
                         term == "home_awayHome"   ~ "Home team (1 / 0)",
                         term == "setterTRUE:post" ~ "Setter × Post‐rule"
                       ),
    β                = sprintf("%.6f", estimate),
    `SE (clustered)` = sprintf("%.6f", std.error),
    `95% CI`         = paste0(
                         sprintf("%.6f", conf.low), " – ",
                         sprintf("%.6f", conf.high)
                       ),
    p                = if_else(
                         p.value < 0.001,
                         "< 0.001",
                         sprintf("%.3f", p.value)
                       )
  )

# 3. Render with kableExtra
kable(
  tbl4,
  booktabs = TRUE,
  caption  = "Table 4. Difference‐in‐Differences Estimates: Setter vs Non‐Setter"
) %>%
  kable_styling(full_width = FALSE)
```

# Weekly mean BHE per point, Setters vs Others (Appendix Figure A2)
```{r setter-non-pre-post, fig.cap='Pre/Post BHE Trends by Position & Season', echo=FALSE}
# Prepare data
player_match$BHE_per_point <- player_match$match_BHE/player_match$match_points

setter_ts4 <- player_match %>%
  left_join(matches %>% select(match_id, match_date), by = "match_id") %>%
  mutate(
    week   = floor_date(match_date, "week"),
    season = factor(season, levels = c("2023-24","2024-25")),
    sex    = factor(sex,    levels = c("W","M"), labels = c("Women","Men"))
  ) %>%
  group_by(sex, season, week, setter) %>%
  summarise(mean_BHE = mean(BHE_per_point, na.rm = TRUE), .groups="drop")

# Plot
ggplot(setter_ts4, aes(x = week, y = mean_BHE, color = setter)) +
  geom_line(size = 1) +
  geom_vline(xintercept = as.Date("2024-08-01"), linetype = "dashed") +
  facet_grid(sex ~ season, scales = "free_x") +
  scale_color_manual(
    values = c("FALSE" = "grey50", "TRUE" = "darkgreen"),
    labels = c("Other", "Setter")
  ) +
  scale_y_continuous(
    labels = label_number(accuracy = 0.0005),  # e.g. 0.000, 0.001, 0.0015
    expand = expansion(mult = c(0, .05))
  ) +
  labs(
    x     = "Week",
    y     = "Mean BHE per Point",
    color = "Position",
    title = "Pre/Post BHE Trends by Position & Season"
    # subtitle removed
  ) +
  theme_minimal() +
  theme(
    strip.text      = element_text(face = "bold", size = 12),
    legend.position = "bottom",
    axis.text.x     = element_text(angle = 45, hjust = 1),
    panel.spacing   = unit(1, "lines")
  )
```

# BHE by Position before/after rule change {-}
## Appendix Figure A2
```{r by-position}
# 1. Total points per season
season_points <- matches %>%
  group_by(season) %>%
  summarize(
    total_points = sum(total_points, na.rm = TRUE)
  )

# 2. Total BHE by season & position
pos_bhe <- players %>%
  group_by(season, P) %>%
  summarize(
    total_BHE = sum(BHE, na.rm = TRUE),
    .groups = "drop"
  )

# 3. Compute BHE per point by joining
pos_rates_by_season <- pos_bhe %>%
  left_join(season_points, by = "season") %>%
  mutate(
    BHE_per_point = total_BHE / total_points
  ) %>%
  arrange(season, P)

bhe_long <- bind_rows(
  # Total BHE
  pos_bhe %>%
    transmute(
      position = P,
      season,
      metric   = "nominal_BHE",
      value    = total_BHE
    ),
  # BHE per Rally
  pos_rates_by_season %>%
    transmute(
      position = P,
      season,
      metric   = "BHE_per_point",
      value    = BHE_per_point
    )
)

# 4. Plot side-by-side bars for each metric
ggplot(bhe_long, aes(x = position, y = value, fill = season)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(
        ~ metric,
        scales   = "free_y",
        labeller = as_labeller(c(
            nominal_BHE   = "Total BHE",
            BHE_per_point = "BHE per Point"    
          ))
      ) +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  labs(
    x     = "Player Position",
    y     = NULL,
    fill  = "Season",
    title = "BHE by position before and after the NCAA rule change"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank()
  )
```

## DiD Estimates by Player Position (2024–25 vs 2023–24) model results
```{r player-positions}
# 1. Create a table of position values aggregated at the match level
player_position <- players %>%
  # bring in match‐level total_points, attendance & season
  left_join(
    matches %>% select(match_id, attendance),
    by = "match_id"
  ) %>%
  # compute controls
  mutate(
    log_attendance = log1p(attendance),
    post           = as.integer(season == "2024-25")
  ) %>%
  # aggregate BHE by match × position
  group_by(
    match_id,
    season,
    P,               # your position code
    post,
    log_attendance
  ) %>%
  summarise(
    match_BHE    = sum(BHE,           na.rm = TRUE),
    match_points = unique(total_points),
    .groups      = "drop"
  ) %>%
  # outcome per point
  mutate(
    BHE_per_point = match_BHE / match_points
  ) %>%
  # make P a factor with setter as reference
  mutate(
    P = relevel(factor(P), ref = "S")
  )

# 3. Fit the DiD model as in earlier model, with P in place of setter
did_position <- feols(
  BHE_per_point ~ post * P + log_attendance |
    match_id + season,
  cluster = ~ match_id,
  data    = player_position
)
summary(did_position)

# 3. Extract only the post:P interaction terms (i.e. the DiD estimates for each position)
did_pos_tbl <- broom::tidy(did_position, conf.int = TRUE) %>%
  filter(str_detect(term, "^post:P")) %>%
  mutate(
    Position    = str_remove(term, "^post:P"),
    Estimate    = round(estimate,  4),
    `Std. Error`= round(std.error,  4),
    `CI Lower`  = round(conf.low,   4),
    `CI Upper`  = round(conf.high,  4),
    `p-value`   = round(p.value,    4)
  ) %>%
  select(Position, Estimate, `Std. Error`, `CI Lower`, `CI Upper`, `p-value`)

# 4. Render as a table
knitr::kable(
  did_pos_tbl,
  caption = "DiD Estimates by Player Position (2024–25 vs 2023–24)",
  align   = c("l", rep("r", 5))
)
```

## Table 5
```{r table-5-position}
# 1.  Create preliminary table
tbl5_rates <- pos_rates_by_season %>%
  # ensure exactly one BHE_per_point per P×season
  group_by(P, season) %>%
  summarise(
    BHE_rate = unique(BHE_per_point),
    .groups = "drop"
  ) %>%
  # now pivot to wide so each P has exactly two columns
  pivot_wider(
    names_from  = season,
    values_from = BHE_rate
  ) %>%
  # rename for your table headers
  rename(
    `2023–24 BHE/pt` = `2023-24`,
    `2024–25 BHE/pt` = `2024-25`
  ) %>%
  # compute the deltas
  mutate(
    `Absolute Δ` = `2024–25 BHE/pt` - `2023–24 BHE/pt`,
    `Relative Δ` = (`Absolute Δ` / `2023–24 BHE/pt`) * 100
  )

# 2.  Pull out post:P interactions
interacts <- tidy(did_position, conf.int = TRUE) %>%
  filter(str_detect(term, "^post:P")) %>%
  transmute(
    P        = str_remove(term, "^post:P"),
    estimate,
    conf.low,
    conf.high,
    p.value
  )

tbl5 <- tbl5_rates %>%
  left_join(interacts, by = "P") %>%
  transmute(
    Position                  = P,
    `2023–24 BHE/pt`          = sprintf("%.6f", `2023–24 BHE/pt`),
    `2024–25 BHE/pt`          = sprintf("%.6f", `2024–25 BHE/pt`),
    `Absolute Δ`              = sprintf("%+.6f", `Absolute Δ`),
    `Relative Δ`              = sprintf("%+.1f%%", `Relative Δ`),
    `DiD vs Setter (Δ BHE/pt, 95% CI)` = if_else(
      Position == "S",
      "—",
      paste0(
        sprintf("%+.5f", estimate),
        " [", sprintf("%+.5f", conf.low), ", ", sprintf("%+.5f", conf.high), "]",
        case_when(
          p.value < 0.001 ~ "***",
          p.value < 0.01  ~ "**",
          p.value < 0.05  ~ "*",
          p.value < 0.10  ~ "†",
          TRUE            ~ ""
        )
      )
    )
  ) %>%
  # Put S first in order
  mutate(
    Position = factor(
      Position,
      levels = c("S", setdiff(unique(Position), "S"))
    ),
    # Relabel only the setter
    Position = fct_recode(Position, "S (ref)" = "S")
  ) %>%
  arrange(Position)

# 3. Render
kable(
  tbl5,
  booktabs = TRUE,
  caption = "Table 5. Change in BHE per Point by Position with DiD vs Setter"
) %>%
  kable_styling(full_width = FALSE)
```


# Home‑Bias Attenuation {-}
## Model results

```{r home-bias}
# 1. Compute home‑away difference per match
match_diff <- teams %>%
  group_by(match_id, season) %>%
  summarise(
    bhe_home = BHE_per_point[home_away == "Home"],
    bhe_away = BHE_per_point[home_away == "Away"],
    log_attendance = unique(log_attendance),
    post = unique(post),
    .groups = "drop"
  ) %>%
  mutate(bhe_diff = bhe_away - bhe_home)

# 2. Fit model
hb_model <- feols(
  bhe_diff ~ post + log_attendance,
  cluster = ~ match_id,
  data    = match_diff
)

# 3. Display results
summary(hb_model)
```

## Table 6
```{r table-6}
# Get the fixest coeftable (with clustered SEs)
fe_summ <- summary(hb_model)              # this has the clustered SEs you saw
ct      <- fe_summ$coeftable            # matrix with Estimate, Std. Error, t, Pr(>|t|)
# Get the confints the same way
cis     <- confint(hb_model, level = 0.95)  # two-column matrix: 2.5% and 97.5%

# Assemble the table
tbl_ha <- tibble(
  Predictor      = c("Intercept", "Post-rule (1 / 0)", "log(Attendance)"),
  beta           = ct[,"Estimate"],
  se_clustered   = ct[,"Std. Error"],
  ci_lower       = cis[,1],
  ci_upper       = cis[,2],
  pval           = ct[,"Pr(>|t|)"]
) %>%
  transmute(
    Predictor,
    beta         = sprintf("%.6f", beta),
    se_clustered = sprintf("%.6f", se_clustered),
    ci95         = paste0(sprintf("%.6f", ci_lower), " – ",
                          sprintf("%.6f", ci_upper)),
    p            = if_else(pval < 0.001, "< 0.001", sprintf("%.3f", pval))
  )

# Render with kableExtra
kable(
  tbl_ha,
  booktabs  = TRUE,
  caption   = "Table 6. Home–Away Difference Model for BHE per Point)",
  col.names = c("Predictor", "β", "SE (clustered)", "95% CI", "p")
) %>%
  kable_styling(full_width = FALSE)
```

# Rally-Length Analysis {-}
## Model results

```{r rally-length}
# 1. Aggregate total attacks per match
rally_match <- teams %>%
  group_by(match_id) %>%
  summarise(total_attacks = sum(TotalAttacks, na.rm = TRUE), .groups = 'drop') %>%
  left_join(
    matches %>% select(match_id, total_points, attendance, season),
    by = "match_id"
  ) %>%
  mutate(
    attacks_per_point = total_attacks / total_points,
    log_attendance    = log1p(attendance),
    post              = as.integer(season == "2024-25")
  )

# 2. Descriptive: mean attacks per point by season
rl_stats <- rally_match %>%
  group_by(season) %>%
  summarise(mean_attacks = mean(attacks_per_point, na.rm = TRUE), .groups = 'drop')
knitr::kable(rl_stats, digits = 3, caption = 'Mean attacks per point by season')

# 3. Model: effect of rule change on rally length
rl_mod <- feols(attacks_per_point ~ post + log_attendance | 0,
                 cluster = ~ match_id,
                 data = rally_match)
summary(rl_mod)
```

## Table 7
```{r table-7}
# Get the fixest coeftable (with clustered SEs)
fe_summ <- summary(rl_mod)              # this has the clustered SEs you saw
ct      <- fe_summ$coeftable            # matrix with Estimate, Std. Error, t, Pr(>|t|)
# Get the confints the same way
cis     <- confint(rl_mod, level = 0.95)  # two-column matrix: 2.5% and 97.5%

# Assemble the table
tbl_rl <- tibble(
  Predictor      = c("Intercept", "Post-rule (1 / 0)", "log(Attendance)"),
  beta           = ct[,"Estimate"],
  se_clustered   = ct[,"Std. Error"],
  ci_lower       = cis[,1],
  ci_upper       = cis[,2],
  pval           = ct[,"Pr(>|t|)"]
) %>%
  transmute(
    Predictor,
    beta         = sprintf("%.6f", beta),
    se_clustered = sprintf("%.6f", se_clustered),
    ci95         = paste0(sprintf("%.6f", ci_lower), " – ",
                          sprintf("%.6f", ci_upper)),
    p            = if_else(pval < 0.001, "< 0.001", sprintf("%.3f", pval))
  )

# Render with kableExtra
kable(
  tbl_rl,
  booktabs  = TRUE,
  caption   = "Table 7. Effect of Rule Change on Rally Length (Attacks per Point)",
  col.names = c("Predictor", "β", "SE (clustered)", "95% CI", "p")
) %>%
  kable_styling(full_width = FALSE)
```

# Supplementary Table S1 {-}
```{r supp-table-s1}
# 1. Summarize matches (matches, sets, rallies)
match_sum <- matches %>%
  group_by(sex, division, season) %>%
  summarise(
    Matches = n_distinct(match_id),
    Sets    = sum(sets_played,   na.rm = TRUE),
    Rallies = sum(total_points,  na.rm = TRUE),
    .groups = "drop"
  )

# 2. Summarize teams (teams count, total BHE)
team_sum <- teams %>%
  group_by(sex, division, season) %>%
  summarise(
    Teams     = n_distinct(Name),
    Total_BHE = sum(BHE, na.rm = TRUE),
    .groups   = "drop"
  )

# 3. Join and label
suppS1 <- team_sum %>%
  left_join(match_sum, by = c("sex","division","season")) %>%
  mutate(
    Sex      = factor(sex, levels = c("W","M"), labels = c("Women","Men")),
    Division = case_when(
      sex=="W" & division=="D1" ~ "I",
      sex=="W" & division=="D2" ~ "II",
      sex=="W" & division=="D3" ~ "III",
      sex=="M" & division=="D1" ~ "I",
      sex=="M" & division=="D3" ~ "III"
    ),
    Season   = season
  ) %>%
  select(Sex, Division, Season, Teams, Matches, Sets, Rallies, Total_BHE)

# 4. Add the overall Total row and order the rows
total_row <- tibble(
  Sex       = "Total",
  Division  = "\u2014",
  Season    = "\u2014",
  Teams     = sum(suppS1$Teams)/2,
  Matches   = sum(suppS1$Matches),
  Sets      = sum(suppS1$Sets),
  Rallies   = sum(suppS1$Rallies),
  Total_BHE = sum(suppS1$Total_BHE)
)

suppS1 <- bind_rows(suppS1, total_row)

suppS1 <- suppS1 %>%
  mutate(
    Teams     = scales::comma(Teams),
    Matches   = scales::comma(Matches),
    Sets      = scales::comma(Sets),
    Rallies   = scales::comma(Rallies),
    Total_BHE = scales::comma(Total_BHE),
    Sex       = factor(Sex, levels = c("Women","Men","Total")),
    Season    = factor(Season, levels = c("2023-24","2024-25","—")),
    Division  = factor(Division, levels = c("I","II","III","—"))
  ) %>%
  arrange(Sex, Season, Division)

# 5. Render
kable(
  suppS1,
  booktabs = TRUE,
  col.names = c("Sex","Division","Season","Teams","Matches","Sets","Rallies","Total BHE"),
  caption = "Supplemental Table S1. Sample counts by sex, division, and season."
) %>%
  kable_styling(full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "top")
```
